Reproducible and programmatic generation of neuroimaging visualizations

### 2022-04-12 manuscript.Rmd
fix1a

---
title: "Reproducible and programmatic generation of neuroimaging visualisations"
author: "Sidhant Chopra, Löic Labache, Winnie Orchard, [...]"
date: "09/03/2022"
output:
  word_document: default
  html_document: default
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
---

@@

Visualising neuroimaging data is one of the primary ways in which we
evaluate data quality, interpret results and communicate findings. These
visualisations are commonly produced using graphic interface-based (GUI)
tools where individual images are opened and display settings are
manually changed until the desired output is reached. The choice to use
GUI-based software's has been driven by a perception of convenience,
flexibility and accessibility. However now there are also many
code-based software packages which are well-documented and don't often
require high-level knowledge of programming, making them more accessible
to the neuroimaging community. These tools are flexible and allow for
the generation of reproducible, high-quality and publication-ready brain
visualisations in only a few lines of code (Figure 1), especially within
the R, Python and Matlab environments.

This article will cover three major advantages of using code-generated
visualisations over GUI based tools: replicability, flexibility and
integration. First, by generating figures using code, you increase the
replicability of your figures for yourself, your collaborators and your
readers. Second, code-based software provides more precise controls over
visualisation settings, while also benefiting from advantages of
programming, such as being able to iteratively generate multiple
figures. Finally, being able to integrate figure generation into your
analysis scripts reduces chances of errors, with more advanced tools now
allowing for seamless weaving together of prose and code. Here, we
review the advantages of learning and using programmatic neuroimaging
visualisations, focusing on benefits to replicability, flexibility and
intergration. We conclude by introducing examples of R-packages which
allow for visualisation of statistics at region-of-interest (ROI),
voxel, vertex and edge level data, followed by a brief discussion of
limitations and functionality gaps in code-based brain visualisation
tools.

@@

In recent years, there have been multiple large-scale efforts
empirically demonstrating the lack of reproducibility of findings from
neuroimaging data (cite poldrack). One common solution proposed for
achieving robust and reliable discoveries has been to encourage
scientific output which can be transparently evaluated and independently
replicated. In practise, what this usually means is openly sharing
detailed methods, materials, code and data. While there is a trend
towards increasing transparency and code sharing in neuroimaging
research (cite), the sharing of code used to generate figures which
include brain renderings and spatial maps has been relatively neglected.
This gap in reproducibility is partly driven by the fact that brain
figures are often created using a manual process of tinkering with
sliders, buttons and overlays on a GUI, concluding with a screenshot and
sometimes beatification in image processing software like Illustrator.
This process inherently makes neuroimaging visualisations difficult, if
not impossible to replicate, even by the authors who created them.

Given that brain figures regularly form the centrepiece of
interpretation within a paper, conference presentations and news
reports, making sure they can be reliably regenerated is crucial for
knowledge generation. By writing and sharing code used to generate brain
visualisations, a direct and traceable link is established between the
underlying data and scientific figure. While the code that produces a
replicable figure doesn't reflect the validity of the scientific finding
or the accuracy of the content figure, it instils transparency and
robustness while demonstrating a desire to further scientific knowledge.
Some would even consider publishing figures which cannot be replicated
as closer to advertising, rather than science (cite blog).

@@

Being able to exactly replicate your figures via code has serious
advantages beyond good open science practises. In particular, the
ability to reprogram inputs (such as statistical maps) and settings
(such as colour schemes and visual orientations) can streamline your
entire scientific workflow. Changing inputs and settings via code allows
for the easy production of multiple figures, such as those resulting
from multiple analyses which require similar visualisations. While some
GUI-based tools do offer command-line access to generate visualisations,
they can lack flexibility to easily generate publication ready figures
and can lose the benefits, such as iteration, provided by your preferred
programming environment. A simple for-loop or copying and pasting the
code with the input and/or setting-of-interest modified can be a
powerful method for exploring visualisation options or creating
multi-panel figures. Likewise, an arduous request from a reviewer to
alter the data processing or analysis becomes less of a burden when the
figure can be re-generated with a few lines of code, as opposed to
re-pasting and re-illustrating them. Having a code-base with
reprogrammable inputs can mean that the generation of visualisations
requires less time, energy and effort than manual GUI-based generation.
This also makes it easier to make figures for each subsequent project
that you work on. Keep in mind that the gains of writing code for your
figures are cumulative, and in addition to improving your programming,
you start to build a code-base for figure generation that you can
continue to reuse and share through your scientific career.

Precise controls via code over visualization settings, such as colour
schemes, legend placement and camera angles, can provide you with much
greater flexibility over visualisations. Nonetheless, part of the appeal
of GUI-based tools is that the pre-sets for such settings can provide a
useful starting point and reduce the decision burden on the user.
However, similar pre-sets are often available in the form of default
settings in most code-based packages, negating the need for the user to
manually enter each and every choice of setting required for making the
image. Most code-based tools also come with detialed documentation, or
even entire papers (cite ggseg and ciftitools) which provide examples of
figures which can be used as starting points or templates for new users.
As the popularity of code-sharing for figure increases, there will be an
cornucopia of templates that can be used as starting points for new
figures.

While brain visualisations are often thought of as the end results of
analyses, they also form a vital part of quality control of imaging
data. Tools to automatically detect artefacts, de-noise the data and
generate derivatives are becoming more robust, but we are not yet at the
stage where visualising the data during processing is no longer need.
Nonetheless, when working with large datasets such as Human Connectome
Project (cite) or UK BioBank (cite), it is unfeasible to use traditional
GUI-based tool to visually check the data. The time it takes to open a
single file and achieve the desired visualisation settings vastly
compounds when working with large datasets. Knowing how to
programmatically generate brain visualisations can allow you to iterate
your visualisation code over each image of a large datasets making
checking the quality of each data processing or analysis step accessible
and achievable. The visual outputs of each iteration can be complied
into accessible documents which can be easily scrolled, with more
advanced usage allowing for the creation of interactive HTML reports,
similar to those created by tools like mriQC (cite). Increasing capacity
to conduct visual quality control on larger datasets can increase the
identification of processing errors and result in more reliable and
valid findings from your data.

@@

Often programming languages such as R, Python and Matlab are used for
the analysis in neuroimaging studies, but the brain visualisation
resulting from these analyses is outsourced to separate GIU-based
visualisation tools such as FSLeyes, Freeview or ITK-snap. Switching
from your analysis environment to a GUI-based visualisation process be a
cumbersome deviation from the scientific workflow. This can make
debugging errors more difficult, as you have to regularly switch program
to visually examine the results of any modifications to analyses. Using
the brain visualisation tools that already exist within your chosen
programming environment can provide instant visual feedback on the
impact of modifications to processing or analyses.

Increasingly popular tools such as R-Markdown, Quarto and Jupyter
Notebook allow for the mixing of prose and code in a single script,
resulting in fully reproducible and publication ready papers. By using
code-based tools available within your preferred environment, brain
visualisations can be directly integrated and embedded within a paper or
report. Some journals that publish neuroimaging studies are moving
towards allowing reproducible manuscripts, including reproducible
figures (cite e-life,f1000,plos), with some even allowing on-demand
re-running of code using cloud services (cite).

Neuroimaging data are often spatially 3D and can have multiple time
points, adding a 4th dimension (e.g. fMRI data). Thus, communicating
findings or evaluating quality using static 2D slices is challenging,
and may not be the best representation of the data, or the
interpretations. While well-curated 3D renderings can help with spatial
localisation (see madan), in the end, static images can only provide an
incomplete representation of the data, and forces researchers to choose
the best angle to show, which often involves compromising one result for
another. An added advantage of some of the code-based tools is that you
can generate 'rich' media like interactive widgets -- figures or
animations, which allow users to zoom, rotate and scroll through slices.
Interacting with a figure in this way can improve scientific
communication of findings. Linking to or even embedding these videos or
interactive figures in papers can greatly enhance the communication of
findings and make your paper more engaging for the reader. Such rich
brain visualisations lend themselves to being embedded or shared on
science communication mediums other than academic papers - such as
presentations, web-sites and social media - all of which can promote the
communication of your research with peers and a reach larger audiences.

### Examples of neuroimaging visualisation packages available in R

The following four sections provide brief examples packages and
functions available in R for visualising voxel, vertex, ROI and
edge-level data. A didactic introduction to these tools is provided in a
accompaning [collection of beginner friendly
R-Notebooks](ttps://github.com/sidchop/neuRo-vis-BrainHack-2021).

```{=tex}
\newpage
\blandscape
```
### Figure 1 [Fix image size ratios]

```{r Figure1 , echo=FALSE, out.height = '80%', out.width = '140%'}
library(patchwork)
library(ggplot2)

@@

patchwork <- figs[[1]]/(figs[[2]]|figs[[3]])|(figs[[6]]/(figs[[8]]|figs[[9]]))/figs[[4]]|figs[[10]]/figs[[11]]

@@

patchwork + plot_annotation(tag_levels = 'A',
                            title = 'Examples of brain visualisations in R')

@@

\
*[Figure 1 - Add Caption and remove panel F]*

@@

##### Voxel-level visualisations.

The `oro.nifti` package allows for the reading, writing, manipulation
and visualisation of voxel-level imaging data of either nifti, analyze
or afni formats. In particular, the `image`, `slice` and `orthographic`
functions allow users to display the desired number of slices in the
desired orientations, while also providing precise control over overlaid
images, colour scales, legend placement and other aesthetic settings
(Fig 1A-B). Smoothing and re-sampling of images is sometimes required
for visualisation purposes, which can be achieved via the `ANTsR`
/`extrantsr` or `fslr` packages.

##### Vertex-level visualisations.

The recently developed `fsbrain` package allows for the visualisation of
vertex-level and atlas-level data which is derieved using FreeSurfer
formats. It contains many flexible functions for visualising
individual-level and group-level measures such as cortical thickness,
volume or surface area (Fig 1E). The package also comes with extensive
guides and documentations to assist users (cite).

@@

##### ROI-level visualisations.

@@

The `ciftiTools` package also allows for surface and voxel-level
visualisation of ROIs for the cortex and sub-cortex respectively (Fig
1E). Flexibly allowing CIFTI format atlases and ROIs to be displayed.

##### Edge-level visualisations.

@@

### Limitations and functionality gaps in R packages for brain visualisation

-   While most of these tools do not require a strong knowledge of
    programming, there can still be a steeper learning curve when
    compared to GUI-based tools.

-   For quick and interactive viewing of single images, sometimes GUI
    tools can be faster (although, fsleyes can be called from python and
    R environments with ease)

-   Visualising data-types such as streamlines from DWI-based
    tractography are still not well represented in code-based tools.

-   Lack of user friendly tools which can render non-standard or
    uncommon atlases or ROIs. Although, with `ciftitools`, this becomes
    easier, as long as atlas follows the CIFTI convention.

-   3D renderings of non-standard ROIs like unique subcortical schemes
    or brain stem are still difficult to do. Can to use complex
    workarounds like ITKsnap + pyVista.

@@

```{r Fig1a, eval=FALSE}
#Figure 1A
#voxel 1

# ADD COMMENTS TO EXPLAIN EACH LINE
library(neurobase)
library(ggplotify)

template <- readnii('data/MNI152_T1_1mm_brain.nii.gz') #load the nifti file into R

effect <- readnii("data/MNI152_effect_size.nii.gz")

#threshold the effect at the desired cut-off
effect[effect<0.3] <- NA

#Plot the effect on the brain template
breaks = seq(.3,.6, by=0.005)

png("data/fig1a.png") 
ortho2(x = template, 
       y = effect, 
       crosshairs = F, 
       bg = "white", 
       NA.x = T, 
       col.y= viridis::viridis(n = 500), 
       xyz = c(70,50,80),
       useRaster = T,
       ycolorbar = TRUE, 
       mfrow = c(1,3)) +
  colorbar(breaks,col =viridis::viridis(n=length(breaks)-1), 
           text.col = "black", labels = TRUE, maxleft = 0.95) 
dev.off()

@@

dti <- readnii('data/sub-001__t1_warped.nii.gz') #load the nifti file into R

@@

png('data/fig1b.png')
overlay(x=robust_window(dti), y=atlas, 
        plot.type = "single", 
        z=c(seq(30,110,10)), 
        col.y =  alpha(rainbow(300), 0.3),
        plane = "axial",  
        useRaster = T, 
        bg = "white",
        NA.x=T, 
        zlim.y =c(1,300))

@@

view_xifti_surface(cii, 
                   hemisphere = "left", 
                   view = "both", 
                   idx = 1,
                   colors = viridis::viridis(n = 100),
                   zlim = c(1,1.5),  
                   legend_embed = T,
                   cex.title = 2)

@@

conmat <- example_weighted_undirected
degree <- rowSums(conmat)[which(rowSums(conmat)!=0)]
fig1h <- brainconn(atlas ="schaefer300_n7", 
                   conmat=conmat, 
                   node.size = degree/2,
                   view="ortho", 
                   edge.color.weighted = T,
                   background.alpha = 0.4, 
                   show.legend = F,
                   edge.color = scale_edge_colour_gradient2(low='yellow', mid='orange', high='red')) 
ggsave(filename = 'data/fig1h.png', plot = fig1h, device = 'png', bg = 'white')

>>>>

---
title: "Reproducible and programmatic generation of neuroimaging visualizations"
author: "Sidhant Chopra, Loïc Labache, Winnie Orchard, Avram Holmes"
date: "09/03/2022"
output:
  pdf_document: default
  html_document: default
  word_document: default
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
---

@@

The visualization of neuroimaging data is one of the primary ways in
which we evaluate data quality, interpret results, and communicate
findings. These visualizations are commonly produced using graphic
interface-based (GUI) tools where individual images are opened and,
within each instance, display settings are manually changed until the
desired output is reached. In large part, the choice to use GUI-based
software has been driven by a perception of convenience, flexibility,
and accessibility. However now there are also many code-based software
packages which are well-documented and do not often require high-level
knowledge of programming, making them more accessible to the
neuroimaging community. These tools are flexible and allow for the
generation of reproducible, high-quality, and publication-ready brain
visualizations in only a few lines of code (Figure 1), especially within
the R, Python and MATLAB environments.

This article will cover three major advantages of using code-generated
visualizations over GUI based tools: replicability, flexibility, and
integration. First, by generating figures using code, you increase the
replicability of your figures for yourself, your collaborators and your
readers. Second, code-based software provides more precise controls over
display settings, while also benefiting from advantages of programming,
such as being able to iteratively and rapidly generate multiple figures.
Finally, being able to integrate figure generation into your analysis
scripts reduces chances of errors and increases the accessibility of
resulting analyses pipeline, with more advanced tools now allowing for
the seamless integration of prose and code. Here, we review the
advantages of learning and using programmatic neuroimaging
visualizations, focusing on benefits to replicability, flexibility and
integration. We conclude by introducing examples of R-packages which
allow for visualization of statistics at region-of-interest (ROI),
voxel, vertex, and edge level data, followed by a brief discussion of
limitations and functionality gaps in code-based brain visualization
tools.

@@

In recent years, there have been multiple large-scale efforts
empirically demonstrating the lack of reproducibility of findings from
neuroimaging data (cite poldrack). One common solution proposed for
achieving robust and reliable discoveries has been to encourage
scientific output which can be transparently evaluated and independently
replicated. In practice, this typically entails openly sharing detailed
methods, materials, code, and data. While there is a trend towards
increasing transparency and code sharing in neuroimaging research
(cite), the sharing of code used to generate figures which include brain
renderings and spatial maps has been relatively neglected. This gap in
reproducibility is partly driven by the fact that brain figures are
often created using a manual process that involves tinkering with
sliders, buttons, and overlays on a GUI, concluding with a screenshot
and sometimes beautification in image processing software like
Illustrator. This process inherently makes neuroimaging visualizations
difficult, if not impossible to replicate, even by the authors
themselves.

Visualization scripts should reflect a core feature of open science.
Given that brain figures regularly form the centerpiece of
interpretation within papers, conference presentations, or news reports,
making sure they can be reliably regenerated is crucial for knowledge
generation. By writing and sharing code used to generate brain
visualizations, a direct and traceable link is established between the
underlying data and the corresponding scientific figure. While the code
that produces a replicable figure doesn't necessarily reflect the
validity of the scientific finding or the accuracy of the associated
content, it allows for reproducibility, instilling transparency and
robustness, while demonstrating a desire to further scientific
knowledge. Some even consider publishing figures which cannot be
replicated as closer to advertising, rather than science (cite blog).

@@

Being able to exactly replicate your figures via code has marked
advantages beyond good open science practices. In particular, the
ability to reprogram inputs (such as statistical maps) and settings
(such as color schemes, thresholds, and visual orientations) can
streamline your entire scientific workflow. Changing inputs and settings
via code allows for the easy production of multiple figures, such as
those resulting from multiple analyses which require similar
visualizations. While some GUI-based tools have historically offered
command-line access to generate replicable visualizations, they can lack
both the flexibility to easily generate publication ready figures and
the benefits, such as iteration, provided by your preferred programming
environment. A simple for-loop or copying and pasting the code with the
input and/or setting-of-interest can be a powerful method for exploring
visualization options or rapidly creating multi-panel figures. Likewise,
an arduous request from a reviewer or collaborator to alter the
pre-processing or analysis becomes less of a burden when the associated
figures can be re-generated with a few lines of code, as opposed to
re-pasting and re-illustrating them manually. Having a code-base with
modifiable inputs can mean that the generation of visualizations
requires less time, energy and effort than image and instance specific
GUI-based generation. This also makes it easier to generate consistent
figures across subsequent projects. Keep in mind that the gains of
writing code for your figures are cumulative, and in addition to
improving your programming, you start to build a code-base for figure
generation that you can continue to reuse and share throughout your
scientific career.

Precise controls via code over visualization settings, such as color
schemes, legend placement and camera angles, can provide you with much
greater flexibility over visualizations. Nonetheless, part of the appeal
of GUI-based tools is that the presets for such settings can provide a
useful starting point and reduce the decision burden on novice users.
However, similar presets are often available in the form of default
settings across most code-based packages, negating the need for the user
to manually enter each and every choice of setting required for creating
an image. Most code-based tools also come with documentation, with
R-packages on the CRAN or **Neuroconductor** repositories having
compulsory requirements for detailed guidance. Recent packages have even
started to include documentation in GitHub repositories, or even entire
papers (cite ggseg and ciftitools) which provide examples of figures
which can be used as starting points or templates for new users. As the
popularity of code-sharing for figure increases, there will be a
cornucopia of templates that can be used as the basis for new figures.

While brain visualizations are often thought of as the end results of
analyses, they also form a vital part of quality control for imaging
data. Tools to automatically detect artefacts, de-noise the data and
generate derivatives are becoming more robust, but we are not yet at the
stage where visualizing the data during processing is no longer
necessary. Nonetheless, when working with large datasets such as Human
Connectome Project (cite) or UK BioBank (cite), it is unfeasible to use
traditional GUI-based tool to visually check the data. The time it takes
to open a single file and achieve the desired visualization settings
vastly compounds when working with large datasets. Knowing how to
programmatically generate brain visualizations can allow you to iterate
your visualization code over each image of a large datasets making
checking the quality of each data processing or analysis step accessible
and achievable. The visual outputs of each iteration can be complied
into accessible documents which can be easily scrolled, with more
advanced usage allowing for the creation of interactive HTML reports,
similar to those created by tools like mriQC (cite). Increasing capacity
to conduct visual quality control on larger datasets can increase the
identification of processing errors and result in more reliable and
valid findings from your data.

@@

Often programming languages such as R, Python and MATLAB are used for
the analysis and non-brain visualizations in neuroimaging studies, but
the brain visualizations resulting from these analyses are outsourced to
separate GIU-based visualization tools such as FSLeyes, Freeview or
ITK-snap. Switching from your analysis environment to a GUI-based
visualization process be a cumbersome deviation from the scientific
workflow. This can make debugging errors more difficult, as you have to
regularly switch program to visually examine the results of any
modifications or adjustments to prior analyses. Using the brain
visualization tools that already exist within your chosen programming
environment can provide instant visual feedback on the impact of
modifications to processing or analyses.

Increasingly popular tools such as R-Markdown, Quarto and Jupyter
Notebook allow for the mixing of prose and code in a single script,
resulting in fully reproducible and publication ready papers. By using
code-based tools available within your preferred environment, brain
visualizations can be directly integrated and embedded within a paper or
report. Some journals that publish neuroimaging studies are moving
towards allowing reproducible manuscripts, including reproducible
figures (cite e-life, f1000,plos), with some even allowing on-demand
re-running of code using cloud services (cite).

Neuroimaging data are often spatially 3D and can have multiple time
points, adding a 4th dimension (e.g., fMRI data). Thus, communicating
findings or evaluating quality using static 2D slices is challenging,
and may not be the best representation of the data, or the
interpretations. While well-curated 3D renderings can help with spatial
localisation (see madan), in the end, static images can only provide an
incomplete representation of the data, and forces researchers to choose
the "best" angle to show, which often involves compromising one result
to emphasize another. An added advantage of some of the code-based tools
is that you can generate 'rich' media like interactive widgets --
figures or animations, which allow users to zoom, rotate and scroll
through slices. Interacting with a figure in this way can improve
scientific communication of findings. Linking to or even embedding these
videos or interactive figures in papers can greatly enhance the
communication of findings and make your paper more engaging for the
reader. Such rich brain visualizations lend themselves to being embedded
or shared on science communication mediums other than academic papers -
such as presentations, websites and social media - all of which can
promote the communication of your research with peers and a reach larger
audiences. This last point is becoming increasingly saliet, as marketing
science on social media has become a core medium for the spreading
discoveries, science communication and even a primary avenue for
obtaining employment for early-career re (Cite?). Overall, public
engagement, whether it is a researcher, a member of a jury or civil
society, is one of the keystones of science, and the images we create
are at the center of the process (Li and Xie, 2020:
10.1177/0022243719881113).

### Examples of neuroimaging visualization packages available in R

The following four sections provide brief examples packages and
functions available in R for visualizing voxel, vertex, ROI and
edge-level data. A didactic introduction to these tools is provided in a
accompanying [collection of beginner friendly
R-Notebooks](ttps://github.com/sidchop/neuRo-vis-BrainHack-2021). This
is not an exhaustive list of packages available for visualizing brain
data in R, rather, the following sections and accompanying notebooks aim
to give the reader a sense of the available options.

```{=tex}
\newpage
\blandscape
```
```{r Figure1 , echo=FALSE, out.height = '80%', out.width = '140%'}
library(patchwork)
library(ggplot2)

@@

patchwork <- figs[[1]]/figs[[2]]|(figs[[6]]/(figs[[8]]|figs[[9]]))/figs[[4]]|(figs[[3]]/figs[[10]]/figs[[11]])

@@

patchwork + plot_annotation(tag_levels = 'A',
                            title = 'Examples of brain visualizations in R')

@@

\
*Figure 1. Examples of brain imaging visualization make in using
different R-packages.* A) Voxel-based statistical map threshold and
overlaid over a T1-weighted template data, with a single axial slice
shown. Made using the the `ortho2` function from the `neurobase`
package. B) A voxel-level cortical parcellation overlaid on T1-weighted
MRI data shown in 9-slice axial orientation. Made using the `overlay`
function from the `neurobase` package. C) A CIFTI format surface atlas
with a corresponding statistic assigned to each regions, with both
hemispheres displayed on a inflated template surface in lateral view.
Made using the `view_xifti_surface` from the `ciftiTools` package. D) A
coronal cross-sectional rendering of X, where a statisitcal alue has
been assigned to each of the X regions.

@@

##### Voxel-level visualizations.

The `oro.nifti` package allows for the reading, writing, manipulation
and visualization of voxel-level imaging data of either nifti, analyze
or afni formats. In particular, the `image`, `slice` and `orthographic`
functions allow users to display the desired number of slices in the
desired orientations, while also providing precise control over overlaid
images, colour scales, legend placement and other aesthetic settings
(Fig 1A-B). Smoothing and re-sampling of images is sometimes required
for visualization purposes, which can be achieved via the `ANTsR`
/`extrantsr` or `fslr` packages.

##### Vertex-level visualizations.

The recently developed `fsbrain` package allows for the visualization of
vertex-level and atlas-level data which is derieved using FreeSurfer
formats. It contains many flexible functions for visualising
individual-level and group-level measures such as cortical thickness,
volume or surface area (Fig 1E). The package also comes with extensive
guides and documentations to assist users (cite).

@@

##### ROI-level visualizations.

@@

The `ciftiTools` package also allows for surface and voxel-level
visualization of ROIs for the cortex and sub-cortex respectively (Fig
1E). Flexibly allowing CIFTI format atlases and ROIs to be displayed.

##### Edge-level visualizations.

@@

### Limitations and functionality gaps in R-packages for brain visualization

Most of these tools introduced above do not require a strong
knowledge of programming, but there is still a steeper learning curve
when compared to GUI-based tools. This is especially true when learning
to making fine adjustments to visual auxiliary such as legend placement,
font size and multi-panel figure positioning, for the purpose of a publication
ready figure. While most tools offer some control over these finer
steps, there are differences between them in feature availability and
usability. Although some interactive widgets with limited functionality 
exist (cite papyr), for quick and interactive viewing of single images, 
sometimes GUI tool scan be faster and for convenience, can sometimes be 
opened directly from within some programming environments.

[surface projections]

Some data types are still not well represented in code-based tools. In 
particular, visualizing streamlines resulting from DWI-based, although 
some progress is being made (see XXX? - dont know). Similarly, visualizing 
custom ROIs as 3D renderings, such as those resuting from voxels within 
subcortex or brain stem, is not yet straight forwared.

@@

```{r Fig1a, eval=FALSE}
#Figure 1A

library(neurobase)
library(ggplotify)

#Load in a nifti (.nii.gz) file of a standardized template so use as a background. 
template <- readnii('data/MNI152_T1_1mm_brain.nii.gz') #load the nifti file into R

#Load in a nifti (.nii.gz) file of the statistic you wanted of overlay, with the same dimensions as the template.  
effect <- readnii("data/MNI152_effect_size.nii.gz")

#Threshold the statistic map so only values about a specified amount are displayed
effect[effect<0.3] <- NA

#Set the breaks/intervas you want on the color bar (e.g. from .3 to .6, by intervals of 0.005)
breaks = seq(.3,.6, by=0.005)

#[Optional] If you want to output the figure as a .png, open a png image device
png("data/fig1a.png")

#Use the ortho2 function from the neurobase package to make the figure
ortho2(x = template,   #Specify the background template
       y = effect,     #Specify the effect you want to overlay on the tempplate
       crosshairs = F, #Remove the cross-hairs
       bg = "white",   #Make the background white 
       NA.x = T,       #Do not display NA values
       col.y= viridis::viridis(n = 500), #Select the colour scale. In this case I have used the viridis scale from the viridis package. 
       xyz = c(70,50,80), #set the x y & z slice you want to visualise
       useRaster = T,   #Sometimes using Raster makes for clearer plots
       ycolorbar = TRUE, #Add a colorbar
       mfrow = c(1,1)) + 
  colorbar(breaks,col = viridis::viridis(n=length(breaks)-1), 
           text.col = "black", labels = TRUE, maxleft = 0.95)  #Set the specifications for the colorbar

dev.off()

@@

#Load in a nifti (.nii.gz) file to use as a background. 
t1 <- readnii('data/sub-001__t1_warped.nii.gz')

@@

png('data/fig1b.png')
overlay(x=robust_window(t1), y=atlas, 
        plot.type = "single", 
        z=c(seq(30,110,10)), 
        col.y =  alpha(rainbow(300), 0.3),
        plane = "axial",  
        useRaster = T, 
        bg = "white",
        NA.x=T, 
        zlim.y =c(1,300))

@@

view_xifti_surface(cii, 
                   hemisphere = "both", 
                   view = "lateral", 
                   idx = 1,
                   colors = viridis::viridis(n = 100),
                   zlim = c(1.2,1.5),  
                   legend_embed = T,
                   cex.title = 2)

@@

conmat <- example_weighted_undirected
degree <- rowSums(conmat)[which(rowSums(conmat)!=0)]
fig1h <- brainconn(atlas ="schaefer300_n7", 
                   conmat=conmat, 
                   node.size = degree/2,
                   view="ortho", 
                   edge.color.weighted = T,
                   background.alpha = 0.4, 
                   show.legend = F,
                   edge.color = scale_edge_colour_gradient2(low='yellow', mid='orange', high='red')) 
ggsave(filename = 'data/fig1h.png', 
       plot = fig1h, 
       device = 'png',
       bg = 'white',
       height = 7, 
       units = "in")
